Notation proposal

### 2011-06-01 analysis.tex
more stuff
git-svn-id: https://svn.ligo.caltech.edu/svn/6th-Floor-Papers/lloid_paper@801 ccdd05d5-56c1-461a-bf98-85f6876523b4

In this section we describe an implementation of the method described in
section \ref{label{sec:method}} suitable for rapid gravitational-wave searches
for compact binary coalescence.  The previous method requires several
computations that can be done before the analysis is actually underway.  Thus
we divide the procedure into two stages 1) an offline planning stage and 2) an
online, low latency filtering stage.  The offline stage can be done before the
analysis is started and updated asychronously, whereas the online stage must keep
up with the detector output and produce search results as rapidly as possible.
In the next two subsections we describe what these stages entail.

@@

The filtering algorithm described in the previous section could be used in a
true sample-in-sample-out realtime system.  This would likely require a
nontrivial integration directly into the data acquisition and storage system of
the gravitational wave observatories.  A slightly more modest goal is to leverage
existing low latency, but not realtime, signal processing infrastructure in
order to implement the algorithm described in the previous section.

We have implemented a prototype of the low latency filtering stage using an
open source signal processing environment called \gstreamer\ \cite{gstreamer}.
\gstreamer\ is a vital component of many Linux systems, providing media
playback, authoring, and streaming on devices from cell phones to desktop
computers to streaming media servers.  Given the similarities of gravitational
wave detector data to audio data it is not surprising that \gstreamer\ is
useful.  In our application, GStreamer excels at queueing, synchronizing,
adding, and bookkeeping many different signals at different sample rates.  It
also provides some useful signal processing elements such as decimators,
\textsc{fir} filters, and interpolators.

\begin{figure}[htbp]
	\includegraphics{figures/lloid-diagram.pdf}
	\caption{Schematic of LLOID pipeline illustrating signal flow.  Circles with arrows represent upsampling \protect\includegraphics{figures/upsample-symbol.pdf} or downsampling \protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus signs represent summing junctions \protect\includegraphics{figures/adder-symbol.pdf}.  Squares \protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample rate decreases from the top of the diagram to the bottom.}
\end{figure}

The filter pipeline consists of six distinct stages.

@@

First, the sample rate of the whitened detector data is reduced to successively
lower sample rates by decimation.  Decimation involves applying an antialiasing
filter to the data, and then downsampling by deleting samples.  We use a
192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample} element.
The detector data is provided at every power-of-two sample rate required by the 
template time sliced described in \ref{FIXME}.

\subsubsection{Time delays}

Rather than implement the time sliced templates as zero padded \fir\ filters as
described in \eqref{FIXME} we instead implement them as shorter, but
appropriately time delayed \fir\ filters that contain only the nonzero samples.
In order to assure that we add the \SNR from the time slices adds appropriate

Each decimated detector data stream becomes the input for one time slice, but
it must be appropriately delayed.

@@

@@

@@

\subsubsection{SNR accumulation}

>>>>

In this section we describe an implementation of the \lloid\ method described
in section \ref{label{sec:method}} suitable for rapid gravitational-wave
searches for compact binary coalescence.  The \lloid\ method requires several
computations that can be done before the analysis is actually underway.  Thus
we divide the procedure into two stages 1) an offline planning stage and 2) an
online, low-latency filtering stage.  The offline stage can be done before the
analysis is started and updated asychronously, whereas the online stage must
keep up with the detector output and produce search results as rapidly as
possible.  In the next two subsections we describe what these stages entail.

@@

The \lloid\ algorithm could be used in a true sample-in-sample-out realtime
system.  However, such a system would likely require integration directly into
the data acquisition and storage system of the gravitational wave
observatories.  A slightly more modest goal is to leverage existing low
latency, but not realtime, signal processing infrastructure in order to
implement the \lloid\ algorithm.  For the near-term this should be a viable 
solution for ultra-low latency searches with $\sim 1s$ intrinsic latency.

We have implemented a prototype of the low latency filtering stage using an
open source signal processing environment called \gstreamer\ \cite{gstreamer}.
\gstreamer\ is a vital component of many Linux systems, providing media
playback, authoring, and streaming on devices from cell phones to desktop
computers to streaming media servers.  Given the similarities of gravitational
wave detector data to audio data it is not surprising that \gstreamer\ is
useful for our purpose. \gstreamer\ also provides some useful stock signal
processing elements such as decimators, filters, and interpolators.  
We have extended the \gstreamer\ framework by developing custom elements
that are available in the \gstlal\ library~\cite{gstlal}.
Figure \ref{fig:pipeline} describes schematically how we implement the \lloid\
algorithm using \gstlal\ and \gstreamer\ components.
%
%
\begin{figure}[htbp]
	\includegraphics{figures/lloid-diagram.pdf}
	\caption{\label{fig:pipeline} Schematic of LLOID pipeline illustrating
signal flow.  Circles with arrows represent upsampling
\protect\includegraphics{figures/upsample-symbol.pdf} or downsampling
\protect\includegraphics{figures/downsample-symbol.pdf}.  Circles with plus
signs represent summing junctions
\protect\includegraphics{figures/adder-symbol.pdf}.  Squares
\protect\includegraphics{figures/fir-symbol.pdf} stand for FIR filters.  Sample
rate decreases from the top of the diagram to the bottom.}
\end{figure}

The filter pipeline drawn in figure \ref{fig:pipeline} consists of several
distinct stages and parallel branches.  We outline the stages below.

@@

First, the sample rate of the whitened detector data is reduced to successively
lower sample rates by decimation.  Decimation involves applying an antialiasing
filter to the data, and then downsampling by deleting samples.  We use a
192-tap \textsc{fir} decimator provided by {\tt Gstreamer's audioresample}
element.  The detector data is provided at every power-of-two sample rate
required by the template time sliced described in \eqref{FIXME}.  These
parallel decimated data streams are fed into parallel fir filter banks in the
next stage.

@@

The \fir\ filter banks are implemented using a \gstlal\ element called {\tt
lal\_firbank}, which produces N channels of filter output from an NxM matrix of
\fir\ filter coefficients.  This element is used in parallel branches in the
pipeline to implement the filtering of the SVD basis filters in each time
slice.  Rather than implement the time sliced templates as zero padded \fir\
filters as described in \eqref{FIXME} we instead implement them as shorter, but
appropriately time advanced \fir\ filters that contain only the nonzero
samples.  Adding the appropriate time offset to the filter output makes up for
the lack of explicit zero padding.  The orthogonal filter outputs must be
reconstructed into the filter output for the underlying physical filter ouputs.

@@

The orthogonal filter output provided by the SVD must be used in linear
combinations in order to obtain the physical filter output for the template
waveforms.  This involves a matrix multiplication.  We use a \gstlal\ element
called {\tt lal\_matrix\_mixer} on the output of each orthogonal filter bank in
order to produce the filter output for the original waveforms in each time
slice.  At this point the time slice filter outputs are not yet at the same
sample rate.

@@

Before we can construct the final \SNR\, the individual time slice filter
outputs must be brought to a common sample rate, or upsampled.  This is done
using the same \gstreamer\ element as was used for decimation, {\tt
audioresample}.

\subsubsection{\SNR accumulation}

In order to accumulate \SNR\ for an early warning, the time slice filter
output is upsampled to the next highest frequency time slice and summed. This
process is iterated until the final \SNR is computed.  However at each stage
of upsampling the summed filter outputs are related to the early warning \SNR
by a simple normalization factor that can be computed in advance. Thus the
\lloid\ algorithm and this implementation leads to a simple early warning
pipeline with little additional work.
